---
title: An R Markdown document converted from "MegDemo1.ipynb"
output: html_document
---

# Start fresh

code for starting fresh?

```{r}
rm(list = ls()) #this is cleaning the memory list
```

# get the data

code for opening a file from a url (file in Excel), name it 'fragility23'?

```{r}
linkGit="https://github.com/Funtamendals-Megana/Demo-1/raw/refs/heads/main/FSI-2023-DOWNLOAD.xlsx"
fragility23=rio::import(file = linkGit)
```

# exploratory commands

find out column names in 'fragility23'?
```{r}
names(x = fragility23)
```

Check the data types in df?
```{r}
str(object = fragility23)
```


show me the first 10 rows?
```{r}
head(x = fragility23,10)
```

show me the last 10 rows?
```{r}
tail(fragility23,10)
```


# transformative commands

## subsetting

keep some columns:
Country, Total,
S1: Demographic Pressures,
P1: State Legitimacy,
E2: Economic Inequality into object 'frag23_sub'
```{r}
grep(pattern = "Country|S1|P1|E2|Total",x = names(fragility23),fixed = F,value = T) # test

keep=grep("Country|S1|P1|E2|Total",names(fragility23),fixed = F,value = T)
frag23_sub=fragility23[,keep]
```

## renaming

rename 'frag23_sub' columns, keep var codes  only
```{r}
names(frag23_sub)[3:5]=c("S1","E2", "P1")
```


## filtering

get me the top ten best countries on the  'E2'  measure from 'frag23_sub'
```{r}
tail(frag23_sub[order(x=-frag23_sub$E2),],10) #option1
tail(frag23_sub[order(x=-frag23_sub$E2),'Country'],10) #option2
```


# some computations

give the statistical description of "frag23_sub"
```{r}
summary(object = frag23_sub)
```


the value of the worst quartile in Total
```{r}
q3_Total=quantile(x = frag23_sub$Total, 
                  probs = 0.75, 
                  na.rm = TRUE)
```


show correlations between "S1","E2", "P1"
```{r}
cor(x=frag23_sub[,-c(1,2)]) #no 'Country', no "Total'
```



## show correlations between "S1","E2", "P1" and their "significance"
#?install.packages("corrtable")

```{r}
corrtable::correlation_matrix(df = frag23_sub[,-c(1,2)])
```


regress P1 and E2 on S1
```{r}
lm(S1~P1+E2,data=frag23_sub)
model <- lm(S1 ~ P1 + E2, data = frag23_sub)
summary(model)
```


# some plotting

give me a plot for the 'P1' variable
```{r}
hist(x = frag23_sub$P1) #base r
```


visual correlation between S1 and E2
```{r}
plot(x=frag23_sub$S1, y=frag23_sub$E2)
```

```{r}
## color points if country is on the worst quartile of Total
frag23_sub$Total>=q3_Total
frag23_sub$worstQt=frag23_sub$Total>=q3_Total

plot(frag23_sub$S1, 
     frag23_sub$E2,pch=20,
     col = as.factor(frag23_sub$worstQt))
```


visual of the regression P1 and E2 on S1
```{r}
library(sjPlot)
plot_models(model)
```

